Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments

### 2016-09-26 papers/2016-Special-Issue/v-special-issue-2016.tex
mainly bib edits but also e2 missing stats in red

\documentclass[%
doc,% saying man here makes it doublespaced; saying doc here instead makes it singlespaced again as supported by the documentclass
%man,
floatsintext,%
apacite%
]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}

%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book

@@

When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}. 
We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances. Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral. 
% MG added/moved out of expt 1
When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}. 
We controlled which side the smaller number appeared on to avoid systematic influences of this effect.

@@

\begin{figure}[htbp]
\fitfigure{images/plotRTandErrorRates-1}
\caption{Experiment 1 results: response times and error rates}
\label{resultse1}
\end{figure}

@@

\begin{table}[htbp]
\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
\label{instre3}
\begin{tabular}{rll}
&crisp&vague\\
\midrule
numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots\\
verbal	&		Choose the square with the fewest dots	&	Choose a square with few dots\\
\midrule
\end{tabular}
\end{table}

@@

\subsubsection{Response times} % E2
Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
A linear mixed model was constructed for the (logged) response times.
The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction. 
\begin{APAenumerate}
	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; 
	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
	% $\beta=.06$, $se=.01$, $t=4.6$, $p<.001$).
	\item [(H2)] Vagueness was disadvantageous in both the \emph{numeric} 
	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
	and \emph{verbal} 
	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
	conditions, offering evidence against hypothesis 2. 
The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format 
	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
	% ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; 
	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
	% ($\beta=.37, se=.07, t=5.1, p<.001$)
	\item [(H4)] There was a significant main effect of which triple of numbers were used in the stimulus 
	($\beta=$, $se=$, $t=$, $p<$). 
	%($\beta=.06, se=.008, t=7.0, p<.001$) 
This effect seems likely to be due to the very fast responses for stimuli 
using arrangements of the smallest numbers of dots (6,15,24), 
which also had the largest difference in ratio of smallest number to largest number in the stimulus, 
suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
\end{APAenumerate}

\subsubsection{Borderline responses} % E2
A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.  
The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}. 
Participant grand mean percentage of borderline selections was $16.6\%$.  
\begin{APAenumerate}
	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, 
	($\beta=??$, $se=??$, $z=??$, $p<??$) % MG ADD STATS
	% $\beta=.79, se=.25, z=3.2, p<.01$).  
\end{APAenumerate}
Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
$\beta=??$, $se=??$, $z=??$, $p<??$). % MG ADD STATS
% $\beta=3.57, se=.26, z=13.6, p<.001$).

\begin{figure}[htbp]
\centering
\includegraphics[width=.49\textwidth]{images/rtplote2.pdf}
\includegraphics[trim = 0mm -30mm 0mm 0mm, clip, width=.49\textwidth]{images/blplote2.pdf}
\caption{Experiment 2 results: response times and proportion of borderline cases}
\label{resultse2}
\end{figure}

@@

% KvD I've expanded the following:
However, once again there is a possibly confounding factor. Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: 
 %as shown by the Table of instructions \ref{instructionse2}, 
%the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements. Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching). % KvD I think we need to say more about this, citing some appropriate papers to support this idea.

@@

>>>>

\documentclass[%
%			% choose man or doc immediately below
doc,			% doc is single spaced
%man,		% man is double spaced
floatsintext,%
apacite%
]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}

%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced

@@

When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}. 
We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances. Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral. 
% MG added/moved out of expt 1
When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}. 
We controlled which side the smaller number appeared on to avoid systematic influences of this effect.

@@

\begin{figure}[htbp]
\fitfigure{images/plotRTandErrorRates-1}
\caption{Experiment 1 results: response times and error rates.}
\label{resultse1}
\end{figure}

@@

\begin{table}[htbp]
\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
\label{instre3}
\begin{tabular}{rll}
&crisp&vague\\
\midrule
numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
\midrule
\end{tabular}
\end{table}

@@

\subsubsection{Response times} % E2
Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
A linear mixed model was constructed for the (logged) response times.
The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction. 
\begin{APAenumerate}
	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	% $\beta=.06$, $se=.01$, $t=4.6$, $p<.001$).
	\item [(H2)] Vagueness was disadvantageous in both the \emph{numeric} 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	and \emph{verbal} 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	conditions, offering evidence against hypothesis 2. 
The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	% ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	% ($\beta=.37, se=.07, t=5.1, p<.001$)
	\item [(H4)] There was a significant main effect of which triple of numbers were used in the stimulus 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). 
	%($\beta=.06, se=.008, t=7.0, p<.001$) 
This effect seems likely to be due to the very fast responses for stimuli 
using arrangements of the smallest numbers of dots (6,15,24), 
which also had the largest difference in ratio of smallest number to largest number in the stimulus, 
suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
\end{APAenumerate}

\subsubsection{Borderline responses} % E2
A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.  
The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}. 
Participants chose the borderline square on $16.6\%$ of trials overall.  
\begin{APAenumerate}
	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, 
	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
	% $\beta=.79, se=.25, z=3.2, p<.01$).  
\end{APAenumerate}
Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). % MG ADD STATS
% $\beta=3.57, se=.26, z=13.6, p<.001$).

\begin{figure}[htbp]
\centering
\includegraphics[width=.49\textwidth]{images/rtplote2.pdf}
\includegraphics[trim = 0mm -30mm 0mm 0mm, clip, width=.49\textwidth]{images/blplote2.pdf}
\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
\label{resultse2}
\end{figure}

@@

% KvD I've expanded the following:
However, once again there is a possibly confounding factor. Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: 
 %as shown by the Table of instructions \ref{instructionse2}, 
%the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements. Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching). % KvD I think we need to say more about this, citing some appropriate papers to support this idea.
\textcolor{red}{insert citation supporting this}

@@

\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}
$\beta=??$, $se=??$, $t=??$, $p<??$
